\inputencoding{latin1}
\HeaderA{modi-package}{Multivariate outlier detection for incomplete survey data}{modi.Rdash.package}
\keyword{package}{modi-package}
%
\begin{Description}\relax
The package modi is a collection of functions for multivariate outlier dedection and imputation. 
The aim is to provide a set of functions which cope with missing values and take sampling weights into account. 
The original functions were 
developed in the EUREDIT project. This work was partially supported by the EU FP5 ICT programme, 
the Swiss Federal Office of Education and Science and the Swiss Federal Statistical Office. 
Subsequent development was in the AMELI project of the EU FP7 SSH Programme and also supported by the 
University of Applied Sciences Northwestern Switzerland FHNW. 
\end{Description}
%
\begin{Details}\relax

\Tabular{ll}{
Package: & modi\\{}
Type: & Package\\{}
Version: & 1.2\\{}
Date: & 2013-05-16\\{}
License: & GPL-2\\{}
LazyLoad: & yes\\{}
}

BACON-EEM algorithm in \code{BEM()}, Epidemic algorithm in \code{EAdet()} and \code{EAimp ()}, Transformed Rank Correlations in \code{TRC()}, Gaussian imputation with MCD in \code{GIMCD()}.  
\end{Details}
%
\begin{Author}\relax
C\bsl{}'edric B\bsl{}'eguin and Beat Hulliger.

Maintainer: Beat Hulliger <beat.hulliger@fhnw.ch>

\end{Author}
%
\begin{References}\relax
B\bsl{}'eguin, C., and Hulliger, B. (2004). Multivariate oulier detection in
incomplete survey data: The epidemic algorithm and transformed
rank correlations. Journal of the Royal Statistical Society, A
167(Part 2.), 275-294.

B\bsl{}'eguin, C. and Hulliger, B. (2008) The BACON-EEM Algorithm for Multivariate Outlier Detection 
in Incomplete Survey Data, \emph{Survey Methodology}, Vol. 34, No. 1, pp. 91-103.
\end{References}
